---
title: "Correlations"
author: "Leslie Jones"
date: "January 29, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo =FALSE)
library(readr)
library(tidyverse)
library(ggplot2)
library(dplyr)
library(lubridate)
library(psych)
library(FactoMineR)
library(ade4)
library(amap)
library(vegan)
library(knitr)
library(reshape2)
library(MASS)
library(data.table)
library(factoextra)
library (colorspace)
library (RColorBrewer)
library(ggplot2)
library(lattice)
library(readr)
library (corrplot)
#install.packages("PerformanceAnalytics")
library("PerformanceAnalytics")
setwd ("C:/Habitat_Diversity/Analysis")
```

## Correlations in Phab data
```{r}
#load data
Phab <- read_csv("C:/Habitat_Diversity/Analysis/Data/Phab_HUClevel_NEW_corr.csv")
#View(Phab)
Phab$cv_elev <- Phab$std_elev/Phab$avg_elev

```
#HUC Geomorphic Correlations
I used indices for acceptable limits of ±2 for skewness(Trochim & Donnelly, 2006; Field, 2000 & 2009; Gravetter & Wallnau, 2014) 
References:
Trochim, W. M., & Donnelly, J. P. (2006). The research methods knowledge base (3rd ed.). Cincinnati, OH:Atomic Dog.
Gravetter, F., & Wallnau, L. (2014). Essentials of statistics for the behavioral sciences (8th ed.). Belmont, CA: Wadsworth.
Field, A. (2000). Discovering statistics using spss for windows. London-Thousand Oaks- New Delhi: Sage publications.
Field, A. (2009). Discovering statistics using SPSS. London: SAGE.
```{r}
shed <- c("dem_cv", "dem_std", "dem_mean", "dem_max", "slope_cv", "slope_std", "slope_mean", "slope_max")
Phab_shed <- Phab[shed]

pairs(Phab_shed)

res <- cor(Phab_shed)
round(res,2)

corrplot(res, type="upper", order="hclust",
         tl.col="black", tl.srt=45)

chart.Correlation(Phab_shed, histogram=TRUE, pch=19)

#test for skewness
geo_huc <- skewness(Phab_shed)
View(geo_huc)
```

#Stream Geomorphic Correlations
```{r}

stream <- c("cv_elev", "std_elev", "avg_elev", "max_elev", "cv_slope", "std_slope", "avg_slope", "max_slope")
Phab_stream <- Phab[stream]

pairs(Phab_stream)

res <- cor(Phab_stream)
round(res,2)

corrplot(res, type="upper", order="hclust",
         tl.col="black", tl.srt=45)

chart.Correlation(Phab_stream, histogram=TRUE, pch=19)

#test for skewness
geo_stream <- skewness(Phab_stream)
View(geo_stream)

```


#Floodplain Correlations
```{r}

flood <- c("flood_elev", "flood_stdarea", "nd_max", "main_sinu", "flood_density", "flood_Tarea")
Phab_flood <- Phab[flood]

pairs(Phab_flood)

res <- cor(Phab_flood)
round(res,2)

corrplot(res, type="upper", order="hclust",
         tl.col="black", tl.srt=45)

chart.Correlation(Phab_flood, histogram=TRUE, pch=19)

#test for skewness
geo_flood <- skewness(Phab_flood)
View(geo_flood)
```

#Hydrology Correlations
```{r}

hydro <- c("avg_l_elev", "std_l_area", "area_lakes", "density_lakes", "n_lakes", "area_glac", "density_glac", "wetland", "wetland_HUC")
Phab_hydro <- Phab[hydro]

pairs(Phab_hydro)

res <- cor(Phab_hydro)
round(res,2)

corrplot(res, type="upper", order="hclust",
         tl.col="black", tl.srt=45)

chart.Correlation(Phab_hydro, histogram=TRUE, pch=19)

#test for skewness
geo_hydro <- skewness(Phab_hydro)
View(geo_hydro)

```

#Landcover Correlations
```{r}

land <- c("forest", "forest_HUC", "shrub_HUC", "bare_HUC", "area_glac", "wetland", "wetland_HUC")
Phab_land <- Phab[land]

pairs(Phab_land)

res <- cor(Phab_land)
round(res,2)

corrplot(res, type="upper", order="hclust",
         tl.col="black", tl.srt=45)

#test for skewness
geo_land <- skewness(Phab_land)
View(geo_land)

```

