---
title: "Species plots"
author: "Leslie Jones"
date: "February 19, 2019"
output: html_document
---

```{r setup,include=FALSE, warning=FALSE, results='hide', message=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(readr)
library(tidyverse)
library(ggplot2)
library(dplyr)
library(lubridate)
library(vegan)
library(knitr)
library(data.table)
library(factoextra)
library (colorspace)
library (RColorBrewer)
library(ggplot2)
library(lattice)
```

## Species Abundance plots for SASAP Regions

```{r, include=FALSE, warning=FALSE, results='hide', message=FALSE}
#load data
species <- read_csv("C:/Habitat_Diversity/Analysis/Data/Abundance_all.csv")
View(species)
```


```{r}
p <- ggplot(data=species, aes(x=SASAP.Region, y=meanCumAnnualCount, fill=SASAP.Region)) + geom_bar(stat = "identity") + facet_wrap(~Species) + theme(axis.text.x = element_text(angle = 90, hjust=1), legend.position = "none")

p + facet_grid(Species ~., scale="free")

```

## Chinook plot without Southeast region since Southeast has much higher Abundance than other regions.
```{r}
chinook <- species %>% filter(Species=="chinook") %>% filter(!SASAP.Region=="Southeast")

ggplot(data=chinook, aes(x=SASAP.Region, y=meanCumAnnualCount, fill=SASAP.Region)) + geom_bar(stat = "identity") +  theme(axis.text.x = element_text(angle = 90, hjust=1), legend.position = "none")
```

# Chum plot without Southeast region since Southeast has much higher Abundance than other regions.
```{r}
chum <- species %>% filter(Species=="chum") %>% filter(!SASAP.Region=="Southeast")

ggplot(data=chum, aes(x=SASAP.Region, y=meanCumAnnualCount, fill=SASAP.Region)) + geom_bar(stat = "identity") +  theme(axis.text.x = element_text(angle = 90, hjust=1), legend.position = "none")
```

##Coho plot without Southeast region since Southeast has much higher Abundance than other regions.
```{r}
coho <- species %>% filter(Species=="coho") %>% filter(!SASAP.Region=="Southeast")

ggplot(data=coho, aes(x=SASAP.Region, y=meanCumAnnualCount, fill=SASAP.Region)) + geom_bar(stat = "identity") +  theme(axis.text.x = element_text(angle = 90, hjust=1), legend.position = "none")
```

##Pink plot without Southeast region since Southeast has much higher Abundance than other regions.
```{r}
pink <- species %>% filter(Species=="pink") %>% filter(!SASAP.Region=="Southeast")

ggplot(data=pink, aes(x=SASAP.Region, y=meanCumAnnualCount, fill=SASAP.Region)) + geom_bar(stat = "identity") +  theme(axis.text.x = element_text(angle = 90, hjust=1), legend.position = "none")
```


##Sockeye plot without Southeast region since Bristol Bay has much higher Abundance than other regions. 
```{r}
sockeye <- species %>% filter(Species=="sockeye") %>% filter(!SASAP.Region=="Bristol Bay")

ggplot(data=sockeye, aes(x=SASAP.Region, y=meanCumAnnualCount, fill=SASAP.Region)) + geom_bar(stat = "identity") +  theme(axis.text.x = element_text(angle = 90, hjust=1), legend.position = "none")
```