---
title: "Regions PHab Analysis"
author: "Leslie Jones"
date: "January 29, 2019"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE, warning=FALSE, results='hide', message=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(readr)
library(tidyverse)
library(ggplot2)
library(dplyr)
library(lubridate)
library(psych)
library(FactoMineR)
library(ade4)
library(amap)
library(vegan)
library(knitr)
library(reshape2)
library(MASS)
library(data.table)
library(factoextra)
library (colorspace)
library (RColorBrewer)
library(ggplot2)
library(lattice)
library(readr)
library (corrplot)
library(moments)
library(BBmisc)
setwd ("C:/Habitat_Diversity/Analysis")

```


```{r, warning=FALSE, results='hide', message=FALSE}
##Data management
#load data
Phab5 <- read_csv("C:/Habitat_Diversity/HUC_data/HUC_habitatData_031819.csv")
#View(Phab5)
huchab <- read_csv("C:/Habitat_Diversity/Analysis/Data/Huc_habitat.csv")
Phab <- left_join (Phab5, huchab, by="id_numeric")

#View(Phab)

#Test for skewness
#drop <- c("region", "X1", "name", "id_numeric", "pink_km", "pink_hab", "coho_km", "coho_hab", "sockeye_km", "sockeye_hab",
         # "chinook_km", "chinook_hab", "chum_km", "chum_hab")
#skewtest <- Phab[ , !(names(Phab) %in% drop)] 
#skewtest2 <- skewtest[complete.cases(skewtest), ]
#View(skewtest2)
#skew <- skewness(skewtest2)
#chart.Correlation(skewtest2, histogram=TRUE, pch=19)
#View (skew)

#log transform skewed variables from 0. Corrlations in Phab Indicators
#Phab$cv_elevT <- log(Phab$cv_elev +1)
#Phab$area_glacT <- log(Phab$area_glac +1)
#Phab$flood_TareaT <- log(Phab$flood_Tarea +1)
#Phab$avg_l_elevT <- log(Phab$avg_l_elev +1)
#Phab$std_l_areaT <- log(Phab$std_l_area +1)
#Phab$area_lakesT <- log(Phab$area_lakes +1)
#Phab$density_lakesT <- log(Phab$density_lakes +1)
#Phab$wetlandT <- log(Phab$wetland +1)

#exclude character variables and redundant for PCA

exclude <- c("name", "region", "id_numeric", "region_id",  "coho_km", "coho_hab",
             "X1", "chinook_km", "chinook_hab", "chum_km", "chum_hab", "pink_km", "pink_hab","sockeye_km", "sockeye_hab")
     
exclude2 <- c("coho_km", "coho_hab", "X1", "chinook_km", "chinook_hab", "chum_km", "chum_hab", "pink_km", "pink_hab",
              "sockeye_km", "sockeye_hab")
Phab2 <- Phab[ , !(names(Phab) %in% exclude2)] 
Phab22 <- Phab2[complete.cases(Phab2), ]
Phab33 <- left_join(Phab22, huchab, by="id_numeric")
#View(Phab33)
```


```{r, include=FALSE, warning=FALSE,  message=FALSE}
##PCA Analysis
Phab9 <- Phab[,!names(Phab) %in% exclude]
Phab99 <- Phab9[complete.cases(Phab9),] 
#chart.Correlation(Phab99, histogram=TRUE, pch=19)
skew <- skewness(Phab99)
#View(skew)
#Phab0 <- sapply(Phab99, function(x)log(x+1))
#Phab0 <- normalize(Phab99, method="standardize", margin=2)
#chart.Correlation(Phab0, histogram=TRUE, pch=19)
#skew2 <- skewness(Phab0)
#View(skew2)
# need to scale and center data
#Phabm<-scale(Phab[,!names(Phab) %in% exclude], scale=TRUE, center=TRUE)
Phabm<-scale(Phab99, scale=TRUE, center=TRUE)
#View(Phabm)

#compute PCA - all data
res.pca <- prcomp(Phabm, scale = FALSE)

#pc1 is a tibble of all pc scores for 170 hucs.
pcscores <- as.tibble(scores(res.pca))
#View(pcscores)
#add huc and region to the pc scores.
pcout <- cbind(Phab33,pcscores)

#View(pcout)
# get variable
var <- get_pca_var(res.pca)
#View(var$cos2)

#get pc scores by var - for objective 1 tables
print(res.pca)

#plot with PC1 scores by var >0.3
#use these to create labels for objective 2 figure
ggplot(pcout, aes(x=PC1, y=elev_avgW)) + geom_point()

ggplot(pcout, aes(x=PC1, y=slope_cvW)) + geom_point()

#export csv with scores by huc and region
#write.csv(pcout, file="C:/Habitat_Diversity/Analysis/Data/PCscores_region.csv")


```


```{r,  warning=FALSE,  message=FALSE}
#vector arrows by phab
#Figure
fviz_pca_var(res.pca,
             col.var = "contrib", # Color by contributions to the PC
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = FALSE     # Avoid text overlapping
)

jpeg(file="C:/Habitat_Diversity/Figures/vectorplot.jpeg")

#scree plot of eigenvalues - variance explained by each PC
fviz_eig(res.pca, addlabels = TRUE, ylim = c(0,40))

screeplot(res.pca, bstick=TRUE, type="lines")

#visualize the cos2 of variables
var <- get_pca_var(res.pca)
corrplot(var$cos2, is.corr = FALSE)

# Total cos2 of variables on Dim.1 and Dim.2
fviz_cos2(res.pca, choice = "var", axes = 1:2)

# Contributions of variables to PC1
fviz_contrib(res.pca, choice = "var", axes = 1, top = 10)
# Contributions of variables to PC2
fviz_contrib(res.pca, choice = "var", axes = 2, top = 10)
# Contributions of variables to PC3
fviz_contrib(res.pca, choice = "var", axes = 3, top = 10)

#graph 
fviz_pca_ind(res.pca,
             col.ind = "cos2", # Color by the quality of representation
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE     # Avoid text overlapping
)

#individuals by region
fviz_pca_ind(res.pca,
             col.ind = Phab22$region, # color by groups
             palette = ("rainbow_hcl"),
             addEllipses =FALSE,
             legend.title = "Groups",
             repel = TRUE
)

#individuals by region with ellipse
fviz_pca_ind(res.pca,
             col.ind = Phab22$region, # color by groups
             palette = ("rainbow_hcl"),
             addEllipses =TRUE,
             legend.title = "Groups",
             repel = TRUE
)


#pc 1 and 2 with points colored and shaped by region, ellipses around regions 
#habillage can be changed to display other factor variables - e.g. variable type.

fviz_pca_ind(res.pca, geom.ind = "point", axes = c(1,2), 
             repel = TRUE, habillage = Phab22$region, addEllipses = FALSE)

fviz_pca_ind(res.pca, geom.ind = "point", axes = c(1,2), 
             repel = TRUE, habillage = Phab22$region, addEllipses = TRUE)

fviz_pca_ind(res.pca, geom.ind = "", axes = c(1,2), 
             repel = TRUE, habillage = Phab22$region, addEllipses = TRUE)

fviz_pca_biplot(res.pca, geom.ind = "point", axes = c(1,2), 
                repel = TRUE, habillage = Phab22$region, addEllipses = TRUE)

fviz_pca_biplot(res.pca, geom.ind = "none", axes = c(1,2), 
                repel = TRUE, habillage = Phab22$region, addEllipses = TRUE)


fviz_pca_ind(res.pca, geom.ind = "", axes = c(1,2), 
             repel = TRUE, habillage = Phab22$region, addEllipses = TRUE, ellipse.type = "confidence")

```

```{r ,  warning=FALSE,  message=FALSE}

#F8766D
#B79F00
#00BA38
#00BFC4
#619CFF
#F564E3

#select ellipsoids


#Arctic Yukon add kuskokwim add kotzebue
fviz_pca_biplot(res.pca, geom.ind = "point", axes = c(1,2), 
                repel = TRUE, habillage = Phab33$region, addEllipses = TRUE,
                palette = c( rep("white",5), "#F8766D", rep("white",7)))


#other similar systems
# Cook Inlet and Bristol Bay copper and southeast and PWS
fviz_pca_biplot(res.pca, geom.ind = "point", axes = c(1,2), 
                repel = TRUE, habillage = Phab33$region, addEllipses = TRUE,
                palette = c( "white", "white", "#B79F00", "white", "#F8766D", "#619CFF", rep("white",4),
                             "yellow3", "#00BFC4", "white"))

# Aleutian,  Norton Sound kodiak chignik
fviz_pca_biplot(res.pca, geom.ind = "point", axes = c(1,2), 
                repel = TRUE, habillage = Phab33$region, addEllipses = TRUE,
                palette = c( "#B79F00", rep("white",2),"#619CFF", rep("white",2), "#00BFC4", 
                             rep("white",2),"#F564E3",rep("white",3)))



```
# Habitat by Species
Totals were created from AWC, Yukon and BC fish distribution data


##Hucs sized by Coho habitat
Standardized habitat totals by HUC.

```{r, warning=FALSE,  message=FALSE}
x1 <- Phab33 %>% pull(coho_hab) 
x11 <- scales::rescale(x1, to=c(1,10))
#View(x11)

fviz_pca_biplot(res.pca, col.var="slategray3", pointsize=(x11), geom.ind = "point", axes = c(1,2), repel = TRUE, habillage = Phab33$region, addEllipses = FALSE)

#set NA hucs to a value so that can be visualized on map
x2 <- Phab33 %>% pull(coho_hab) 
x2 [x2>0] <- 1 
x4 <- x2 %>% replace_na(0)
View(x4)

x3 <- Phab33 %>% pull(coho_hab) 
x31 <- scales::rescale(x3, to=c(1,10)) %>% replace_na(1)

View(x31)

fviz_pca_biplot(res.pca, col.var="slategray3", pointsize=x31, geom.ind = "point", axes = c(1,2), repel = TRUE, habillage = x4, addEllipses = FALSE, invisible = "quali")

#fviz_pca_biplot(res.pca, col.var="slategray3", pointsize=3, geom.ind = "point", axes = c(1,2), repel = TRUE, habillage = x4, #addEllipses = FALSE)
                
```

##Hucs sized by Chinook habitat
Standardized habitat totals by HUC.

```{r, warning=FALSE,  message=FALSE}
x1 <- Phab33 %>% pull(chinook_hab) 
x11 <- scales::rescale(x1, to=c(1,10))
#View(x11)

fviz_pca_biplot(res.pca, col.var="slategray3", pointsize=(x11), geom.ind = "point", axes = c(1,2), repel = TRUE, habillage = Phab33$region, addEllipses = FALSE)

#set NA hucs to a value so that can be visualized on map
x2 <- Phab33 %>% pull(chinook_hab) %>%replace_na(0)
x2 [x2>0] <- 1 
#View(x2)

x3 <- Phab33 %>% pull(chinook_hab) 
x31 <- scales::rescale(x3, to=c(1,10)) %>% replace_na(1)

#View(x31)

fviz_pca_biplot(res.pca, col.var="slategray3", pointsize=x31, geom.ind = "point", col.ind=as.factor(x2), axes = c(1,2), repel = TRUE, habillage = "none", addEllipses = FALSE, invisible = "quali")

#fviz_pca_biplot(res.pca, col.var="slategray3", pointsize=3, geom.ind = "point", axes = c(1,2), repel = TRUE, habillage = x2, #addEllipses = FALSE)
```




##Hucs sized by Sockeye habitat
Standardized habitat totals by HUC.

```{r, warning=FALSE,  message=FALSE}
x1 <- Phab33 %>% pull(sockeye_hab) 
x11 <- scales::rescale(x1, to=c(1,10))
#View(x11)

fviz_pca_biplot(res.pca, col.var="slategray3", pointsize=(x11), geom.ind = "point", axes = c(1,2), repel = TRUE, habillage = Phab33$region, addEllipses = FALSE)

#set NA hucs to a value so that can be visualized on map
x2 <- Phab33 %>% pull(sockeye_hab) %>%replace_na(0)
x2 [x2>0] <- 1 
#View(x2)

x3 <- Phab33 %>% pull(sockeye_hab) 
x31 <- scales::rescale(x3, to=c(1,10)) %>% replace_na(1)

#View(x31)

fviz_pca_biplot(res.pca, col.var="slategray3", pointsize=x31, geom.ind = "point", axes = c(1,2), repel = TRUE, habillage = x2, addEllipses = FALSE, invisible = "quali")

#fviz_pca_biplot(res.pca, col.var="slategray3", pointsize=3, geom.ind = "point", axes = c(1,2), repel = TRUE, habillage = x2, #addEllipses = FALSE)
     
```

##Hucs sized by Pink habitat
Standardized habitat totals by HUC.

```{r, warning=FALSE,  message=FALSE}
x1 <- Phab33 %>% pull(pink_hab) 
x11 <- scales::rescale(x1, to=c(1,10))
#View(x11)

fviz_pca_biplot(res.pca, col.var="slategray3", pointsize=(x11), geom.ind = "point", axes = c(1,2), repel = TRUE, habillage = Phab33$region, addEllipses = FALSE)

#set NA hucs to a value so that can be visualized on map
x2 <- Phab33 %>% pull(pink_hab) %>%replace_na(0)
x2 [x2>0] <- 1 
#View(x2)

x3 <- Phab33 %>% pull(pink_hab) 
x31 <- scales::rescale(x3, to=c(1,10)) %>% replace_na(1)

#View(x31)

fviz_pca_biplot(res.pca, col.var="slategray3", pointsize=x31, geom.ind = "point", axes = c(1,2), repel = TRUE, habillage = x2, addEllipses = FALSE, invisible = "quali")

#fviz_pca_biplot(res.pca, col.var="slategray3", pointsize=3, geom.ind = "point", axes = c(1,2), repel = TRUE, habillage = x2, #addEllipses = FALSE)
```     

##Hucs sized by Chum habitat
Standardized habitat totals by HUC.

```{r, warning=FALSE,  message=FALSE}
y1 <- Phab33 %>% pull(chum_hab) 
y11 <- scales::rescale(x1, to=c(1,10)) 
#View(y11)

fviz_pca_biplot(res.pca, col.var="slategray3", pointsize=y11, geom.ind = "point", axes = c(1,2), repel = TRUE, habillage = Phab33$region, addEllipses = FALSE)

#set NA hucs to a value so that can be visualized on map
x2 <- Phab33 %>% pull(chum_hab) %>%replace_na(0)
x2 [x2>0] <- 1 
#View(x2)

x3 <- Phab33 %>% pull(chum_hab) 
x31 <- scales::rescale(x3, to=c(1,10)) %>% replace_na(1)

#View(x31)

fviz_pca_biplot(res.pca, col.var="slategray3", pointsize=x31, geom.ind = "point", axes = c(1,2), repel = TRUE, habillage = x2, addEllipses = FALSE, invisible = "quali")

#fviz_pca_biplot(res.pca, col.var="slategray3", pointsize=3, geom.ind = "point", axes = c(1,2), repel = TRUE, habillage = x2, #addEllipses = FALSE)
```